---
title: "Tuto_package"
author: "Amélie Picard and Baptiste Lanoue"
date: "07/12/2020"
output: 
  md_document:
  html_document:
    df_print: paged
---


This vignette gives a high level overview on how to use the clustyanaly R package
```{r setup}
library(clustyanaly)

```
Use clustyanaly index
------------------------

We will show you how we can use internal index from the package 


```{r}
data("iris")
df <- iris
head(df, 3)
```
We will use iris data and we predict a kmean with cluster = 5
parameter must be a data frame and vector

```{r}
fit <- kmeans(scale(df[, -5]), 5)
clust <- fit$cluster
clust <- as.vector(clust)
df <- as.data.frame(scale(df[, -5]))
```
Silhouette index
-----------------

You can evaluate the silhouette index for each point of the data frame, cluster and general 
Value are beetween -1 and 1
```{r}
silhouette(df,clust)
```

You can also choose a distance 
```{r}
silhouette(df,clust,"manhattan")
```
Davies bouldin index
---------------------

You can evaluate the Davies_bouldin index
the lower the value, the better is the clustering
```{r}
davies_bouldin(df,clust)
```
You can also choose a distance 
```{r}
davies_bouldin(df,clust,"maximum")
```
Dunn index
-----------

You can evaluate the Dunn index
the higher the value, the better is the clustering
You have different evaluation of Dunn in function of the distance beetwenn cluster(mean,max,min)
```{r}
dunn(df,clust)
```
You can also choose a distance 
```{r}
dunn(df,clust,"canberra")
```
Variance index
-----------------

You can evaluate different variance index like calinski_harabasz coefficient, Ball_hall coefficient, Hartigan coefficient and Xu coefficient.
```{r}
variance_index(df,clust)

```
Example with Davies Bouldin index
-----------------------------------

We can find the best number cluster for our model with Davies Bouldin index for example
```{r}
fit <- kmeans(scale(df[, -5]), 5)
clust <- fit$cluster
clust <- as.vector(clust)
df <- as.data.frame(scale(df[, -5]))
DB_id <- c()
for (k in 2:20){
  fit <- kmeans(scale(df[, -5]), k)
  clust <- fit$cluster
  clust <- as.vector(clust)
  DB_id <- append(DB_id ,davies_bouldin(df,clust))
}
cluster_number <- c(2:20)
plot(cluster_number,DB_id)  
```

Tutorial for univariate and multivariate characterization
-----------------------------------------------------------

We are going to graphically represent three datasets. We take for this tutorial a data frame with quantitative variable "mtcars", a data frame with qualitative variable "canine" and finally a data frame with both type of data "small_apb".

Load data :

The data frame "canines" and "small_apb" have already a column with the class value of individuals. It's the last column of the data frame. 

Load the canines data : 
```{r}
#group variable :
groupe_canines<-canines[8]
#analysis variables
df_canines<-canines[c(-1,-8)]
rownames(df_canines)<-canines[,1]
```

Load the small_apb data : 
```{r}
#group variable :
apb_groupe<-small_apb[,length(small_apb)]
#analysis variables
apb<-small_apb[c(-1,-length(small_apb))]
```

The data frame "mtcars" does not have columns representing groups of individuals. We are therefore going to make a hierarchical ascending classification (hac) to create groups.

Load the mtcars data:
```{r}
#center and reduce data
mtcars.cr <- scale(mtcars,center=T,scale=T)
#calculating of distance
d.mtcars <- dist(mtcars.cr)
#hierarchical ascending classification with method ward
cah.ward2 <- hclust(d.mtcars,method="ward.D2")

#group variable with 4 class:
groupes.cah2 <- cutree(cah.ward2,k=4)
#analysis variables : mtcars

```



Creation of type 'ClusteringData' :
------------------------------------

Thanks to the function "data_manager()", we can create an object of class 'ClusteringData'. This object allows you to simply use the other functions accessible from the package.

```{r}
#for canine data
canine_objet<-data_manager(df_canines,groupe_canines)
#for mtcars data
cars<-data_manager(mtcars,groupes.cah2)
#for small_apb data
apb_objet<-data_manager(apb,apb_groupe)
```

Display data graphs univariate
-------------------------------

Display data graphs univariate by data type (quantitative or qualitative variables). If the data has both types of data, all graphs are displayed, else only the corresponding graphs are displayed.

For qualitative data :
```{r}
#for canine data
get_graph_car_univ(canine_objet)
```

For quantitative data : 
```{r}
#for mtcars data
get_graph_car_univ(cars)
```

For mixed data : 
```{r}
#For small_apb data
get_graph_car_univ(apb_objet)
```

For a better analysis of quantitative variables, it's possible to display matching graphics

```{r}
#For small_apb data
get_graph_car_univ(apb_objet,choice_graph = "quanti")

```


Display data graphs multivariate
----------------------------------

Display data graphs multivariate according to factor analysis performed.

For qualitative data :
```{r}
mca_graph(canine_objet)
```

For quantitative data : 
```{r}
pca_graph(cars)
```

For mixed data : 
```{r}
famd_graph(apb_objet)
```

Some features are modifiable like, which graphic it is possible to display, and which dimension of factor analysis is to represent or the colors of the graphics. 
Note : that the color vector must have at least the same number of colors as the data has classes.

Example : 
----------

We want to see the graphs, for mixed data, containing information on individuals. And for AFMD dimensions 1 and 3. 

```{r}
famd_graph(apb_objet,axes = c(1,3),graph = c("ind","indpvar"))
```


